---
title: "Introduction to assessor"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{assessor}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# assessor

The goal of assessor is to provide assessment tools for regression models with discrete and semicontinuous outcomes proposed in [Yang (2023)](https://arxiv.org/pdf/2308.15596.pdf). It calculates the double probability 
integral transform (DPIT) residuals, constructs QQ plots of residuals and the ordered curve for assessing mean 
structures.


## General function arguments
In `assessor`, there are functions for assessing DPIT residuals for the various type of models: discrete, zero-inflated, and semi-continuous outcome regression models. `resid_disc()`, `resid_zeroinfl()`, `resid_semiconti()` and `resid_2pm()` are functions evaluating DPIT residuals. 
They share the same arguments:

* `model`: Each function supports different types of model objects. Check below which model objects are applicable.

* `plot`:If you set `plot=T`, a function will return QQ-plot drawn with DPIT residuals, or vice versa. 
Some users may only need the QQ-plot. In that case, `qqresid()` only returns the QQ-plot without returning the DPIT residuals.

* `scale`: You can choose the scale of the residuals among `normal` and `uniform` scales. The sample quantiles of the residuals are plotted against the theoretical quantiles of a standard normal distribution under the normal scale, and against the theoretical quantiles of a uniform (0,1) distribution under the uniform scale. The default scale is `normal`

## Real data example
The `solder` data, extracted from the `faraway` package, lends itself to fitting through either Poisson regression or Negative Binomial regression. The selection between these models depends on the underlying distribution characteristics of the data. To validate the distribution assumption, a comprehensive approach involves assessing the DPIT residuals and scrutinizing the QQ plot derived from these residuals.

```{r example, fig.width= 7, fig.height=3}
library(assessor)
library(faraway) 
library(MASS)
data("solder") # from faraway package


## Negative Binomial
modpnb <- glm.nb(skips~.,data=solder)
modp <- glm(skips~.,famil=poisson(link="log"),data=solder)

## QQ-plot
par(mfrow=c(1,3))
poi.resid <- resid_disc(modp,plot = T)
norm.resid <- resid_disc(modpnb,plot = T, scale = "normal")
unif.resid <- resid_disc(modpnb,plot = T, scale = "uniform")
```
The figure above depicts three versions of QQ plots. The second and third plots essentially represent the same graph, corresponding to the QQ plot for Negative Binomial regression (`modpnb`). Both plots align closely along the diagonal line, suggesting that the assumption of Negative Binomial distribution appears appropriate. The only distinction between them lies in the `scale` parameter; the middle panel displays the plot in normal scale, with the $x$ and $y$ axes spanning values larger than 1, while the right panel limits the axes to a range of 0 to 1.

Conversely, the left panel corresponds to the QQ plot from Poisson regression (`modp`). In this plot, a slight deviation from the diagonal line is observed. Consequently, we can infer that the outcome variable, `skips`, is more likely to follow a Negative Binomial distribution rather than a Poisson distribution. This inference is drawn based on the examination of QQ plots generated by DPIT residuals, highlighting the suitability of the Negative Binomial model for capturing the distributional characteristics of the data.




