---
title: "Introduction to assessor"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{assessor}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# assessor

The goal of assessor is to provide assessment tools for regression models with discrete and semicontinuous outcomes proposed in [Yang (2023)](https://arxiv.org/pdf/2308.15596.pdf). It calculates the double probability 
integral transform (DPIT) residuals, constructs QQ plots of residuals and the ordered curve for assessing mean 
structures.


## General function arguments
In `assessor`, there are functions for assessing DPIT residuals for the various type of models: discrete, zero-inflated, and semi-continuous outcome regression models. `resid_disc()`, `resid_zeroinfl()`, `resid_semiconti()` and `resid_2pm()` are functions evaluating DPIT residuals. 
They share the same arguments:

* `model`: Each function supports different types of model objects. Check below which model objects are applicable.

* `plot`:If you set `plot=T`, a function will return QQ-plot drawn with DPIT residuals, or vice versa. 
Some users may only need the QQ-plot. In that case, `qqresid()` only returns the QQ-plot without returning the DPIT residuals.

* `scale`: You can choose the scale of the residuals among `normal` and `uniform` scales. The sample quantiles of the residuals are plotted against the theoretical quantiles of a standard normal distribution under the normal scale, and against the theoretical quantiles of a uniform (0,1) distribution under the uniform scale. The default scale is `normal`

## Example
This tab explains 

```{r example, fig.width= 7, fig.height=3}
library(assessor)
library(faraway) 
library(MASS)
data("solder") # from faraway package


## Negative Binomial
modpnb <- glm.nb(skips~.,data=solder)
modp <- glm(skips~.,famil=poisson(link="log"),data=solder)

## QQ-plot
par(mfrow=c(1,3))
poi.resid <- resid_disc(modp,plot = T)
norm.resid <- resid_disc(modpnb,plot = T, scale = "normal")
unif.resid <- resid_disc(modpnb,plot = T, scale = "uniform")
```



