---
title: "Introduction to assessor"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to assessor}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# assessor

The goal of assessor is to provide assessment tools for regression models with discrete and semicontinuous outcomes proposed in Yang's papers on [discrete outcomes (2024)](https://doi.org/10.1080/10618600.2024.2303336) and [semicontinuous outcomes (2024)](https://doi.org/10.1093/biomtc/ujae007). It calculates the double probability
integral transform (DPIT) residuals, constructs QQ plots of residuals for model diagnostics, and constructs the ordered curve for assessing mean 
structures.


## General workflow
In `assessor`, there are functions for calculating DPIT residuals for the various type of models: discrete, zero-inflated, and semi-continuous outcome regression models. `dpit()` and `dpit_2pm()` are functions evaluating DPIT residuals for discrete, zero-inflated discrete, semicontinuous, and two-part outcomes (the former functions `resid_disc()`, `resid_semiconti()`, and `resid_zeroinfl()` have been unified under `dpit()`, and `resid_2pm()` has been renamed to `dpit_2pm()`).
Both functions return a `dpit` object. After supplying a fitted model to `dpit()`, or model objects or fitted probabilities to `dpit_2pm()`, use the returned object's methods:

* `model`: `dpit()` supports certain types of model objects. `dpit_2pm()` accepts models for the zero and positive parts or their fitted probabilities. Check below on which model objects are applicable.

* `residuals()`: Use this accessor to obtain the numeric DPIT residual values from the returned object.

* `summary()`: Use this method to summarize the DPIT residual values on the selected scale.

* `plot()`: Use this method to draw a QQ-plot of the DPIT residuals from the returned object.

* `scale`: In `residuals()`, `summary()`, and `plot()`, you can choose the scale of the residuals among `normal` and `uniform` scales. The sample quantiles of the residuals are plotted against the theoretical quantiles of a standard normal distribution under the normal scale, and against the theoretical quantiles of a uniform (0,1) distribution under the uniform scale. The default scale is `normal`.

## Real data example
The `bballHR` data included in `assessor` contain annual home run counts and batted-ball characteristics for Major League Baseball players. The count outcome can be fitted using either Poisson or negative binomial regression. To assess the distributional assumption, we compare the DPIT residuals and their QQ plots.

```{r example, fig.width= 7, fig.height=3}
library(assessor)
library(MASS)
data("bballHR")


## Negative Binomial
model_formula <- HR ~ mean_exit_velo + mean_launch_angle + offset(log(AB))
modpnb <- glm.nb(model_formula, data = bballHR)
modp <- glm(model_formula, family = poisson(link = "log"), data = bballHR)

## QQ-plot
par(mfrow=c(1,3))
poi.dpit <- dpit(modp)
poi.resid <- residuals(poi.dpit, scale = "normal")
plot(poi.dpit, scale = "normal")

norm.dpit <- dpit(modpnb)
norm.resid <- residuals(norm.dpit, scale = "normal")
plot(norm.dpit, scale = "normal")

unif.dpit <- dpit(modpnb)
unif.resid <- residuals(unif.dpit, scale = "uniform")
plot(unif.dpit, scale = "uniform")
```

The left panel of the figure above presents the QQ plot of the DPIT residuals when we use a Poisson GLM (`modp`). The residuals display a pronounced S-shaped pattern, hinting at potential overdispersion. 

The second and third panels present the of DPIT residuals of the negative binomial regression (`modpnb`) on different scales. Both plots align closely along the diagonal line, suggesting that the assumption of negative binomial distribution appears appropriate. The only distinction between them lies in the `scale` parameter of `plot()`; the middle panel displays the plot on the normal scale, with the $x$ and $y$ axes spanning beyond (0,1), while the right panel limits the axes to a range of 0 to 1.
