---
title: "assesor"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{assesor}
  %\VignetteEngine{knitr::knitr}
  %\VignetteEncoding{UTF-8}{inputenc}
---

## Introduction
Assessor is a package that provides tools for assessing the assumptions of regression models with discrete, zero-inflated, and semi-continuous outcomes. Users can calculate the double probability integral transform (DPIT) residuals and draw QQ-plots via the DPIT residuals. 

### `resid_disc()`
`resid_disc()` is used for calculating the DPIT residuals of regression models with discrete outcomes and drawing QQ-plots. The suitable distribution is as follows: 

* Poisson, `glm(formula, family=poisson(link="log"))` 
* Binary, `glm(, family=binomial(link="logit"))` 
* Negative Binomial, `MASS::glm.nb()` 
* Ordinal, `MASS::polr()` 

And then, the arguments of the function is as follows:

* `model`: model object 
* `plot`:
* `scale`: 

```{r discrete outcomes poi, fig.align = 'center', eval=FALSE, warning=FALSE}
library(assessor)
library(faraway)
library(MASS)

## Solder data
data("solder")
par(mfrow=c(1,3))
## Poisson with main effects
modp <- glm(skips~.,famil=poisson(link="log"),data=solder)
resid.modp <- resid_disc(modp,plot = TRUE)

## NB with main effects
modpnb <- glm.nb(skips~.,data=solder)
resid.modpnb <- resid_disc(modpnb,plot = TRUE,scale = "normal")

## NB with interactions 
modpnb1 <- stepAIC(glm.nb(skips ~ (Opening +Solder + Mask + PadType + Panel)^2 , data=solder),trace = FALSE)
resid.modpnb1 <- resid_disc(modpnb1,plot = TRUE,scale = "normal")

## Ordered curve
par(mfrow=c(1,2))
ord_curve(modpnb,modpnb$fitted.values)
ord_curve(modpnb1,modpnb1$fitted.values)
```


```{r disc nb}

```

### `resid_zeroinfl()`
```{r zeroinflated outcomes}

```

### `resid_semiconti()`
```{r}

```


### `qqresid()`
```{r plotting}

```

### `ord_curve()`
```{r }

```


