---
title: "`resid_2pm()`: Two part regression"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


### Two part models {.tabset .tabset-pills}
The input arguments for the `resid_2pm()` function differ from those of other functions in `assessor` package. Specifically, users can utilize this function with either models or Probability Integral Transform (PIT) as input.

For instance, in evaluating the distribution assumption of a two-part model that combines logistic and gamma regression, you should provide the logistic regression model object as the argument for `model0` and the gamma regression model object for `model1`. We recommend utilizing the model input when assessing a gamma+logistic two-part model. Alternatively, users can directly use PIT as input if their two-part model does not consist of a gamma+logistic combination. In such cases, users should first calculate the PIT and then input the results into `part0` and `part1`, respectively.

This function accommodates two combinations: either `model0` or `part0` in conjunction with either `model1` or `part1`. Note that it is essential to specify the `y` (outcome) values in the function arguments.

#### Models as input
```{r model as input}
library(assessor)
library(MASS)
n <- 500
beta10 <- 1
beta11 <- -2
beta12 <- -1
beta13 <- -1
beta14 <- -1
beta15 <- -2
x11 <- rnorm(n)
x12 <- rbinom(n, size = 1, prob = 0.4)

p1 <- 1 / (1 + exp(-(beta10 + x11 * beta11 + x12 * beta12)))
lambda1 <- exp(beta13 + beta14 * x11 + beta15 * x12)
y2 <- rgamma(n, scale = lambda1 / 2, shape = 2)
y <- rep(0, n)
u <- runif(n, 0, 1)
ind1 <- which(u >= p1)
y[ind1] <- y2[ind1]
```

```{r 2pm 2, fig.align='center'}
# models as input
mgamma <- glm(y[ind1] ~ x11[ind1] + x12[ind1], family = Gamma(link = "log")) # Gamma regression
m10 <- glm(y == 0 ~ x12 + x11, family = binomial(link = "logit")) # logistic regression

resid.models <- resid_2pm(model0 = m10, model1 = mgamma, y = y)
```


#### PIT as input
```{r pit as input,fig.align='center'}
library(assessor)
library(MASS)
n <- 500
beta10 <- 1
beta11 <- -2
beta12 <- -1
beta13 <- -1
beta14 <- -1
beta15 <- -2
x11 <- rnorm(n)
x12 <- rbinom(n, size = 1, prob = 0.4)

p1 <- 1 / (1 + exp(-(beta10 + x11 * beta11 + x12 * beta12)))
lambda1 <- exp(beta13 + beta14 * x11 + beta15 * x12)
y2 <- rgamma(n, scale = lambda1 / 2, shape = 2)
y <- rep(0, n)
u <- runif(n, 0, 1)
ind1 <- which(u >= p1)
y[ind1] <- y2[ind1]

# PIT as input
mgamma <- glm(y[ind1] ~ x11[ind1] + x12[ind1], family = Gamma(link = "log")) # gamma regression
m10 <- glm(y == 0 ~ x12 + x11, family = binomial(link = "logit")) # logistic regression

cdfgamma <- pgamma(y[ind1],
  scale = mgamma$fitted.values * gamma.dispersion(mgamma),
  shape = 1 / gamma.dispersion(mgamma)
)
p1f <- m10$fitted.values

resid.pit <- resid_2pm(part0= p1f, part1 = cdfgamma, y = y)
```

