---
title: "Discrete"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


### Discrete outcome regression models {.tabset .tabset-pills}
`resid_disc()` is used for calculating the DPIT residuals for regression models with discrete outcomes and drawing corresponding QQ-plots. The suitable model objects are as follows: 

* Negative Binomial, `MASS::glm.nb()` 
* Poisson, `glm(formula, family=poisson(link="log"))` 
* Binary, `glm(, family=binomial(link="logit"))` 
* Ordinal, `MASS::polr()`


#### Negative binomial
An appropriate example of the usage of the `resid_disc()` function is in the context of Negative binomial regression. We can do this using the code below. 
```{r discrete outcomes nb}
library(assessor)
library(MASS)
n <- 500
set.seed(1234)
## Negative Binomial example
# Covariates
x1 <- rnorm(n)
x2 <- rbinom(n, 1, 0.7)

### Parameters
beta0 <- -2
beta1 <- 2
beta2 <- 1
size1 <- 2
lambda1 <- exp(beta0 + beta1 * x1 + beta2 * x2)

# generate outcomes
y <- rnbinom(n, mu = lambda1, size = size1)
```
Description here

```{r nb res, fig.align='center', fig.width=10}
par(mfrow=c(1,2))
# True model
model1 <- glm.nb(y ~ x1 + x2)
resd1 <- resid_disc(model1, plot = TRUE, scale = "uniform")

# Overdispersion
model2 <- glm(y ~ x1 + x2, family = poisson(link = "log"))
resd2 <- resid_disc(model2, plot = TRUE, scale = "normal")
```
The interpretation is here


#### Poisson
```{r poisson}
## Poisson example
n <- 500
set.seed(1234)
# Covariates
x1 <- rnorm(n)
x2 <- rbinom(n, 1, 0.7)
# Coefficients
beta0 <- -2
beta1 <- 2
beta2 <- 1
lambda1 <- exp(beta0 + beta1 * x1 + beta2 * x2)
y <- rpois(n, lambda1)
```
Description here

```{r poisson 2, fig.align='center', fig.width=10}
par(mfrow=c(1,2))
# True model
poismodel1 <- glm(y ~ x1 + x2, family = poisson(link = "log"))
resid1 <- resid_disc(poismodel1, plot = TRUE)

# Enlarge three outcomes
y <- rpois(n, lambda1) + c(rep(0, (n - 3)), c(10, 15, 20))
poismodel2 <- glm(y ~ x1 + x2, family = poisson(link = "log"))
resid2 <- resid_disc(poismodel2, plot = TRUE)
```
Interpretation here




#### Binary
```{r bin 1}
## Binary example
n <- 500
set.seed(1234)
# Covariates
x1 <- rnorm(n, 1, 1)
x2 <- rbinom(n, 1, 0.7)
# Coefficients
beta0 <- -5
beta1 <- 2
beta2 <- 1
beta3 <- 3
q1 <- 1 / (1 + exp(beta0 + beta1 * x1 + beta2 * x2 + beta3 * x1 * x2))
y1 <- rbinom(n, size = 1, prob = 1 - q1)
```

```{r bin2,  fig.align='center', fig.width=10} 
par(mfrow=c(1,2))
# True model
model01 <- glm(y1 ~ x1 * x2, family = binomial(link = "logit"))
resid1 <- resid_disc(model01, plot = TRUE)

# Missing covariates
model02 <- glm(y1 ~ x1, family = binomial(link = "logit"))
resid2 <- resid_disc(model02, plot = TRUE)
```

#### Ordinal
```{r ordinal 1}
## Ordinal example
n <- 500
set.seed(1234)
# Covariates
x1 <- rnorm(n, mean = 2)
# Coefficient
beta1 <- 3

# True model
p0 <- plogis(1, location = beta1 * x1)
p1 <- plogis(4, location = beta1 * x1) - p0
p2 <- 1 - p0 - p1
genemult <- function(p) {
  rmultinom(1, size = 1, prob = c(p[1], p[2], p[3]))
}
test <- apply(cbind(p0, p1, p2), 1, genemult)
y1 <- rep(0, n)
y1[which(test[1, ] == 1)] <- 0
y1[which(test[2, ] == 1)] <- 1
y1[which(test[3, ] == 1)] <- 2
multimodel <- polr(as.factor(y1) ~ x1, method = "logistic")
```
True model here
```{r ordinal 2}
## Non-Proportionality
n <- 500
set.seed(1234)
x1 <- rnorm(n, mean = 2)
beta1 <- 3
beta2 <- 1
p0 <- plogis(1, location = beta1 * x1)
p1 <- plogis(4, location = beta2 * x1) - p0
p2 <- 1 - p0 - p1
genemult <- function(p) {
  rmultinom(1, size = 1, prob = c(p[1], p[2], p[3]))
}
test <- apply(cbind(p0, p1, p2), 1, genemult)
y1 <- rep(0, n)
y1[which(test[1, ] == 1)] <- 0
y1[which(test[2, ] == 1)] <- 1
y1[which(test[3, ] == 1)] <- 2
multimodel2 <- polr(as.factor(y1) ~ x1, method = "logistic")
```
Non proportinality here

```{r ordinal 3, fig.align='center', fig.width=10}
par(mfrow=c(1,2))
resid1 <- resid_disc(multimodel, plot = TRUE)
resid2 <- resid_disc(multimodel2, plot = TRUE)
```
