---
title: "Quasi Emprical residuals functions"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

### Discrete outcome regression {.tabset .tabset-pills}

`resid_quasi()` generates QQ-plots for regression models with discrete outcomes, employing quasi-empirical residual distribution functions. It's tailored to assess model assumptions in GLMs featuring binary, ordinal, Poisson, negative binomial, zero-inflated Poisson, and zero-inflated negative binomial outcomes. Unlike typical functions in `assessor` package, `resid_quasi()` exclusively focuses on plotting QQ-plots and does not compute DPIT residuals.

* Negative binomial, `MASS::glm.nb()` 
* Poisson, `glm(formula, family=poisson(link="log"))` 
* Binary, `glm(formula, family=binomial(link="logit"))` 
* Ordinal, `MASS::polr()`
* Zero-Inflated Poisson, `pscl::zeroinfl(dist = "poisson")` 
* Zero-Inflated negative binomial, `pscl::zeroinfl(dist = "negbin")`

The tabs below explain how to interpret the QQ-plots generated by resid_quasi() in Poisson and Zero-inflated Poisson examples, respectively. 

#### Poisson example
We simulate a Poisson random variable using covariates $X_1$ and $X_2$.
The true mean of $Y$ is intricately connected to both $X_1$ and $X_2$, as expressed in the ensuing relationship:
$$
Y \sim \text{Poisson}(\lambda = \exp(\beta_0 + \beta_1 x_1 + \beta_2 x_2)),
$$
where $\beta_0=-2,~\beta_1=2,~\beta_2=1$.


```{r poisson}
library(assessor)
## Poisson example
n <- 500
set.seed(1234)
# Covariates
x1 <- rnorm(n)
x2 <- rbinom(n, 1, 0.7)

# Coefficients
beta0 <- -2
beta1 <- 2
beta2 <- 1
lambda1 <- exp(beta0 + beta1 * x1 + beta2 * x2)
y <- rpois(n, lambda1)
```

```{r poisson 2, fig.align='center', fig.width=10}
# True model
poismodel1 <- glm(y ~ x1 + x2, family = poisson(link = "log"))
resid_quasi(poismodel1)
```
The figure presented above represents the results of `poismodel1`, which is a GLM fitted with a Poisson distribution. In this context, the variable `y` follows a Poisson distribution as defined in the model. Our expectation was that the QQ-plot would exhibit alignment with diagonal lines, as this alignment indicates conformity with the assumptions of the Poisson distribution for discrete outcome regression.

As anticipated, the result indeed demonstrates a well-aligned QQ plot, closely following the diagonal line. This alignment is indicative of the correctness of our model assumption regarding the distribution of the outcome variable. In simpler terms, it suggests that our model is appropriately capturing the characteristics of the data, specifically the discrete nature of the outcomes, as dictated by the Poisson distribution.

#### Zero-Inflation Poisson
We generate simulated data using a zero-inflated Poisson model. The probability of excess zeros is modeled using $\mathrm{logit}(p_0) = \beta_{00} + \beta_{10}X_1$, while the Poisson component has a mean of $\lambda = \exp(\beta_0 +\beta_1X_1 +\beta_2X_2)$. Here, $X_1$ follows a normal distribution with mean 0 and standard deviation 1, and $X_2$ is a binary variable with a probability of 1 set to 0.7. The parameter values are set to $( \beta_{00} ,\beta_{10}, \beta_0, \beta_1, \beta_2) = (-2, 2, -2, 2, 1)$.

```{r zeroinfl poi, message=FALSE}
## Zero-Inflated Poisson
library(assessor)
library(pscl)
n <- 500
set.seed(1234)
# Covariates
x1 <- rnorm(n)
x2 <- rbinom(n, 1, 0.7)
# Coefficients
beta0 <- 0.5
beta1 <- 2
beta2 <- 1
beta00 <- -2
beta10 <- 2

# Mean of Poisson part
lambda1 <- exp(beta0 + beta1 * x1 + beta2 * x2)
# Excess zero probability
p0 <- 1 / (1 + exp(-(beta00 + beta10 * x1)))
## simulate outcomes
y0 <- rbinom(n, size = 1, prob = 1 - p0)
y1 <- rpois(n, lambda1)
y <- ifelse(y0 == 0, 0, y1)
```

```{r 2, fig.align='center', fig.width=10}
par(mfrow=c(1,2))
## True model
modelzero1 <- zeroinfl(y ~ x1 + x2 | x1, dist = "poisson", link = "logit")
resid_quasi(modelzero1)

## Zero inflation
modelzero2 <- glm(y ~ x1 + x2, family = poisson(link = "log"))
resid_quasi(modelzero2)
```
The QQ plots shown above correspond to `modelzero1` and `modelzero2`. Given that the true distribution of `y` follows a zero-inflated Poisson distribution, we expect to see deviations from the diagonal line in the QQ plot of `modelzero2`. As anticipated, the QQ plot on the left closely follows the diagonal line. However, in the right panel, both the left and right tails of the QQ plot for `modelzero2` deviate from the diagonal line, indicating that the assumption of a Poisson distribution may not be accurate.

The observed differences in the QQ plot of `modelzero2` suggest that the assumption of a Poisson distribution is not well-supported by the data. This underscores the importance of considering alternative distributional assumptions, such as the zero-inflated Poisson distribution, which may better capture the characteristics of the simulated data.
