---
title: "Semicontinuous"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

### Semicontinuous outcome regression models {.tabset .tabset-pills}
`resid_semiconti()` is used for calculating the DPIT residuals for regression models with semicontinuous outcomes and drawing corresponding QQ-plots. Specifically, a Tobit regression and a Tweedie regression model are suitable models for `resid_semiconti()`. The suitable model objects are as follows: 

* Tweedie, `glm(family= tweedie())`
* Tobit(VGAM), `VGAM::vglm()` 
* Tobit(AER), `AER::tobit()` 

#### Tweedie 
```{r tweedie1}
## Tweedie model
library(assessor)
library(tweedie)
library(statmod)
n <- 500
x11 <- rnorm(n)
x12 <- rnorm(n)
beta0 <- 5
beta1 <- 1
beta2 <- 1
lambda1 <- exp(beta0 + beta1 * x11 + beta2 * x12)
y1 <- rtweedie(n, mu = lambda1, xi = 1.6, phi = 10)
```
We simulate `y1` to follow Tweedie distribution depending on covariates `x11` and `x12`. 

```{r tweedie2, fig.align='center', fig.width=10}
# True model
model1 <-
  glm(y1 ~ x11 + x12,
    family = tweedie(var.power = 1.6, link.power = 0)
  )

# missing covariate
model2 <-  glm(y1 ~ x11 ,
    family = tweedie(var.power = 1.6, link.power = 0)
  )

par(mfrow=c(1,2))
resid1 <- resid_semiconti(model1)
resid2 <- resid_semiconti(model2)
```
In constructing `model2`, the intentional omission of the covariate `x12` was aimed at facilitating a direct comparison with the true model, `model1`. As expected, the QQ plot in the right panel corresponding to `model2` exhibits a substantial deviation from the diagonal line, attributing this deviation to the deliberate omission. Conversely, the left panel shows a closer alignment along the diagonal, implying a better fit when including both covariates, `x11` and `x12`. This outcome strongly suggests that incorporating all covariates results in a more appropriate and improved model.


#### Tobit
`resid_semiconti()` function supports calculating DPIT residuals for a Tobit regression from both `VGAM::vglm` and `AER::tobit` packages.

```{r tobit vgam1, message=FALSE}
## Tobit regression model
library(VGAM)
beta13 <- 1
beta14 <- -3
beta15 <- 3

set.seed(123)
x11 <- runif(n)
x12 <- runif(n)
lambda1 <- beta13 + beta14 * x11 + beta15 * x12
sd0 <- 0.3
yun <- rnorm(n, mean = lambda1, sd = sd0)
y <- ifelse(yun >= 0, yun, 0)
```
The variable `y` has been generated to conform to a Tobit model depending on the covariates `x11` and `x12`. These variables will be employed as inputs for Tobit regression analyses  provided by the `VGAM` and `AER` packages.

```{r tobit vgam2, fig.align='center', fig.width=10}
# Using VGAM package
# True model
fit1 <- vglm(formula = y ~ x11 + x12, tobit(Upper = Inf, Lower = 0, lmu = "identitylink"))
# Missing covariate
fit1miss <- vglm(formula = y ~ x11, tobit(Upper = Inf, Lower = 0, lmu = "identitylink"))

par(mfrow=c(1,2))
resid1 <- resid_semiconti(fit1, plot = TRUE)
resid2 <- resid_semiconti(fit1miss, plot = TRUE)
```
The model `fit1miss`, corresponding to the right QQ plot, intentionally omits the covariate x12. This omission leads to a deviation from the diagonal line in the QQ plot, observed in the right panel. Thus, it can be interpreted as the model misspecification.  In contrast, the left panel, corresponding to the model including all covariates, aligns closely along the diagonal line. 

In the left panel, there is one point that deviates from the diagonal line, indicating the presence of an outlier.

```{r tobit aer, fig.align='center', fig.width=10, message=FALSE}
# Using AER package
library(AER)
# True model
fit2 <- tobit(y ~ x11 + x12, left = 0, right = Inf, dist = "gaussian")
# Missing covariate
par(mfrow=c(1,2))
fit2miss <- tobit(y ~ x11, left = 0, right = Inf, dist = "gaussian")
reisd1 <- resid_semiconti(fit2, plot = TRUE)
resid2 <- resid_semiconti(fit2miss, plot = TRUE)
```
The interpretation remains the same as the `VGAM` example. Note that the results from the `AER` are exactly the same as those from the `VGAM` example.
