---
title: "Semicontinuous"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

### Semicontinuous outcome regression models {.tabset .tabset-pills}
`resid_semiconti()` is used for calculating the DPIT residuals for regression models with semicontinuous outcomes and drawing corresponding QQ-plots. Specifically, a Tobit regression and a Tweedie regression model are suitable models for `resid_semiconti()`. The suitable model objects are as follows: 

* Tweedie, `glm(family= tweedie())`
* Tobit(VGAM), `VGAM::vglm()` 
* Tobit(AER), `AER::tobit()` 

#### Tweedie 
```{r tweedie1, eval=FALSE}
## Tweedie model
library(assessor)
library(tweedie)
library(statmod)
n <- 500
x11 <- rnorm(n)
x12 <- rnorm(n)
beta0 <- 5
beta1 <- 1
beta2 <- 1
lambda1 <- exp(beta0 + beta1 * x11 + beta2 * x12)
y1 <- rtweedie(n, mu = lambda1, xi = 1.6, phi = 10)
xi.vec <- seq(1.1, 1.9, by = 0.1)
# Choose parameter p
out.model <-
  tweedie.profile(y1 ~ x11 + x12,
    xi.vec = xi.vec,
    do.plot = FALSE,
    verbose = FALSE
  )
```

```{r tweedie2, fig.align='center', eval=FALSE}
# True model
model1 <-
  glm(y1 ~ x11 + x12,
    family = tweedie(var.power = out.model$xi.max, link.power = 0)
  )
resid1 <- resid_semiconti(model1)
```


#### Tobit
```{r tobit vgam1, eval=FALSE}
## Tobit regression model
library(VGAM)
beta13 <- 1
beta14 <- -3
beta15 <- 3

set.seed(1234)
x11 <- runif(n)
x12 <- runif(n)
lambda1 <- beta13 + beta14 * x11 + beta15 * x12
sd0 <- 0.3
yun <- rnorm(n, mean = lambda1, sd = sd0)
y <- ifelse(yun >= 0, yun, 0)
```

```{r tobit vgam2, fig.align='center', fig.width=10, eval=FALSE}
# Using VGAM package
# True model
fit1 <- vglm(formula = y ~ x11 + x12, tobit(Upper = Inf, Lower = 0, lmu = "identitylink"))
# Missing covariate
fit1miss <- vglm(formula = y ~ x11, tobit(Upper = Inf, Lower = 0, lmu = "identitylink"))

par(mfrow=c(1,2))
resid1 <- resid_semiconti(fit1, plot = TRUE)
resid2 <- resid_semiconti(fit1miss, plot = TRUE)
```

```{r tobit aer, fig.align='center', fig.width=10, eval=FALSE}
# Using AER package
library(AER)
# True model
fit2 <- tobit(y ~ x11 + x12, left = 0, right = Inf, dist = "gaussian")
# Missing covariate
par(mfrow=c(1,2))
fit2miss <- tobit(y ~ x11, left = 0, right = Inf, dist = "gaussian")
reisd1 <- resid_semiconti(fit2, plot = TRUE)
resid2 <- resid_semiconti(fit2miss, plot = TRUE)
```
